---
title: 'Blog 3: Running an ANOVA'
author: "Michael O'Donnell"
date: "September 28, 2020"
output:
  pdf_document: default
  html_document: default
---

In Blog Two I explored the Multiple Linear Regression.
Now, in this Blog I will dive into Multiple Regression with a categorical variable, an ANOVA. This will take measure a categorical variable's effect on the response variable.

To look at an ANOVA in R, let's use help
```{r help}
help(aov)
```

Now, start by loading a dataset
This dataset contains data for all NBA teams from 2014-2018
```{r load nba data}
nbaData <- read.csv("data/nba_data.csv")
colnames(nbaData)[1] <- "Team"

head(nbaData, 3)
```

For this analysis, we will test whether the Season (2014-2018) has any impact on Points in the Paint (PointsInPaint).
Y: PointsInPaint
X1: SeasonType

Run an ANOVA
(first variable in Y (response))
```{r build ANOVA}
model1 <- aov(PointsInPaint ~ Season, nbaData)

summary(model1)
```

From the above summary, Season has a statistically significant impact on Points in the Paint. The p-value is far below the significance value of 0.05 and the F value is large.

To view the descriptive statistics by Seaon, we can use the psych library:
```{r}
library(psych)
describeBy(nbaData$PointsInPaint, nbaData$Season)
```

To visualize the data above, we can use ggplot to graph the Points in the Paint by Season
```{r}
library(ggplot2)

ggplot(nbaData,aes(y=PointsInPaint, x=Season))+
  stat_summary(fun="mean", geom="bar",position="dodge")+
  stat_summary(fun.data = mean_se, geom = "errorbar", position="dodge",width=.8)
```
