---
title: 'Blog 2: dropping predictors in Multiple Linear Regression'
author: "Michael O'Donnell"
date: "September 27, 2020"
output:
  pdf_document: default
  html_document: default
---

In Blog One I explored the metrics of Simple Linear Regression.
Now, in this Blog I will dive into Multiple Linear Regression and determine how to drop predictor variables to achieve a higher .

To look at a linear model in R, let's use help
```{r help}
help(lm)
```

Now, start by loading a dataset
This dataset contains regular season data for all NBA teams from 2014-2018
```{r load nba data}
nbaData <- read.csv("data/nba_data.csv")
colnames(nbaData)[1] <- "Team"

head(nbaData, 3)
```

For this analysis, we model the relationship between Points Variables and WinPercentage
Y: WinPercentage
X1: Pts
X2: FastBreakPts
X3: PointsInPaint
X4: PointsOffTO
X5: SecondChancePts
X6: ShootingFoulsDrawnPG

Build simple linear regression model
(first variable in Y (response))
```{r build SLR model}
model1 <- lm(WinPercentage ~ Pts + FastBreakPts + PointsInPaint +
              PointsOffTO + SecondChancePTS + ShootingFoulsDrawnPG, nbaData)
model1
```

Now, lets see a summary of the model
```{r}
summary(model1)
```

From the summary above, we can see that only 3 of the 6 predictor variables are less than our significance value of 0.05 (shown with *'s next to predictors)
Also, the Adjusted R-squared of 0.318 is lower than we want.

So, to determine how attain the best predictors for the model, we use the step function: 
```{r step1}
step(model1, test = "F")
```

From the step function, we can see the best AIC score is achieved with only 3 variables.
Thus, lets create a second model with only the suggested variables:
```{r create second model}
model2 <- lm( WinPercentage ~ Pts + PointsInPaint + PointsOffTO, 
    data = nbaData)

summary(model2)
```

Now, the F-statistic is higher and the adjusted r-squared is a little higher.
The p-value is below the significance value, but the model still doesn't look great.

To diagnose some of the problems of the model, we can use diagnotic plots:
```{r}
plot(model2)
```

In future blog posts, we will dig into the diagnostic plots. For now, we know how to drop predictors from a model.